---
title: "ProjML"
author: "Sunil"
date: "September 12, 2019"
output: html_document
---

### ML Week 4 - Project
Read Input files
```{r }

fileURL = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
destFile <- tempfile()
download.file(fileURL, destFile, method = "curl")

train = read.csv(destFile, na.strings=c("NA","","#DIV/0!"))

fileURL = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
destFile <- tempfile()
download.file(fileURL, destFile, method = "curl")

test = read.csv(destFile, na.strings=c("NA","","#DIV/0!"))


```

## Read Training and Test data

Data - Classe description
Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E)

Source : http://groupware.les.inf.puc-rio.br/har

## Submission
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.


#Install packages
```{r }
library("caret")
library(randomForest)
library(e1071)
library(ggplot2)
set.seed(7777)

```

## Data


Using CART algorithm

```{r }
#head(train)
#head(test)
inTrain = createDataPartition(train$classe, p = 0.6, list = FALSE)
training = train[ inTrain,]
testing = train[-inTrain,]

#colnames(training)
training= training[,-c(1:7)]
testing=testing[,-c(1:7)]
test=test[,-c(1:7)]
training= training[,colSums(is.na(training)) == 0]
testing= testing[,colSums(is.na(testing)) == 0]
test = test[,colSums(is.na(testing)) == 0]
model= train(classe ~ . ,data=training, method ="rpart", na.action=na.omit,trControl = trainControl("cv", number = 04),tuneLength = 10)

print(model, digits=4)

#predTest <- predict(model, newdata=testing)



#summary(model)
par(xpd = NA)
plot(model$finalModel)
text(model$finalModel, digits = 3)
predictTest = predict(model, newdata=testing,na.action=na.pass)
#predictTest
testing$classe = predictTest

varImp(model)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


The 20 variables that are important based on percentage for this analysis is present in the output varImp(model)


 
 
# Validation
```{r }
print(predict(model, newdata=test))


```